---
title: "Untitled"
author: "Josip Domazet"
date: '20 08 2019 '
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages

```{r}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(corrplot)
library(psych)
library(leaflet)
library(naniar) 
library(knitr)
library(kableExtra)
library(plotly)
library(cowplot)
library(caret)
library(randomForest)
library(rpart.plot)
library(rpart)
```

## Knowing our data

```{r cars}
dat <- read.csv("data/countries.csv", encoding="UTF-8")
glimpse(dat)
dat$GDP.per.Capita <-  as.numeric(gsub('[$,]', '', dat$GDP.per.Capita))

```
We can see that 19 of the columns are numeric types. Column GDP.per.Capita had to be converted to 
_double_ type. Other columns are factors. We will ommit _Data.Quality_ because we will not use it anywhere.

### Checking for missing data

```{r}
missing_data <- dat %>% summarise_all(funs(sum(is.na(.))/n()))
missing_data <- gather(missing_data, key = "variables", value = "percent_missing")
missing_data <- missing_data[missing_data$percent_missing > 0.0, ] 
ggplot(missing_data, aes(x = reorder(variables, percent_missing), y = percent_missing)) +
geom_bar(stat = "identity", fill = "red", aes(color = I('white')), size = 0.3)+
xlab('variables')+
coord_flip()+ 
theme_fivethirtyeight() +
  ggtitle("Missing Data By Columns",
          subtitle = "HDI has the highest percentage")

```

```{r}
table1_dat <- dat[is.na(dat$HDI), c(1,2)]
rownames(table1_dat) <- NULL
table1_dat %>% kable() %>%  kable_styling(bootstrap_options = "striped", full_width = T)
```
They are all pretty much small countries. 
Hence, we can ommit them from our dataset.


```{r}
dat$Data.Quality <- NULL
```

## Visualising all numeric columns

```{r}
multi.hist(dat[,sapply(dat, is.numeric)])
```

Most of distrubutions are skewed. 
We will inspect Total Ecological Footprint more detaily.

```{r}

dat %>% ggplot(aes(x = Total.Ecological.Footprint)) +
  geom_histogram(bins = 30, aes(y = ..density..), fill = "purple") + 
  geom_density(alpha = 0.2, fill = "purple") +
  theme_fivethirtyeight() +
  ggtitle("Total per Capita Footprint") +# scale_x_continuous(trans = 'log10') + 
  theme(axis.title = element_text(), axis.title.x = element_text()) +# ovo je napravljeno zato što 538
  geom_vline(xintercept = mean(dat$Total.Ecological.Footprint), size = 1.75, linetype = 3) +
  annotate("text", x = 7, y = 0.35, label = "Average per Capita Footprint is 3.32")


```

## Biggest polluters

```{r}
dat %>% arrange(desc(Total.Ecological.Footprint)) %>% select(Total.Ecological.Footprint, Country) %>% head(n = 15) %>% dplyr::rename(Footprint_per_Person = Total.Ecological.Footprint) %>% 
kable() %>% 
    kable_styling("striped", full_width = F) %>%
  row_spec(1:5, bold = T, color = "white", background = "#D7261E")

```


## Visualising correlation

```{r}
k <- dat[, sapply(dat, is.numeric)]
k <- k[complete.cases(k), ]
korelacija <- cor(k) # izlaz je matrica
corrplot(korelacija, method = "color", tl.cex = 0.825)
```
```{r}
k2 <- dat[, sapply(dat, is.numeric)]
k2 <- k2[complete.cases(k2), ]
korelacija2 <- cor(k2, method = "spearman") # izlaz je matrica
corrplot(korelacija2, method = "color", tl.cex = 0.825)
```

We can see that the results are different, 
so we can conclude that the linear connections is not the best relationship.


```{r pressure, echo=FALSE}
dat %>% group_by(Region) %>% summarise(Total = sum(Total.Ecological.Footprint)) %>% 
  ggplot(aes(Region, Total, fill = Region)) + geom_histogram(stat = "identity") +
  theme_fivethirtyeight() +
  ggtitle("Total per Capita Pollution by Region") +
  theme(axis.text.x = element_text(angle = 20, vjust = 0.65))

```


```{r}
multi.hist(dat[,sapply(dat, is.numeric)])

```



## Linear regression

### Economical Featurs
```{r}

linear_model1 <- lm(Total.Ecological.Footprint ~ HDI, data = dat)
summary(linear_model1)

linear_model2 <- lm(Total.Ecological.Footprint ~ I(exp(HDI)), data = dat)
summary(linear_model2)

linear_model3 <- lm(Total.Ecological.Footprint ~ HDI + I(HDI**2), data = dat)
summary(linear_model3)

slr <- ggplot(dat, aes(HDI, Total.Ecological.Footprint)) +
  geom_point(aes(text = Country)) +
  geom_smooth(method= "lm", color = "red", linetype = 1, se=F) +
  geom_smooth(method= "lm", formula = (y ~ x + I(x**2)), color = "blue", linetype = 2, se=F) +
  ggtitle("Simple Linear Regression",
          subtitle = "Model With Quadratic Term does much better") 

ggplotly(slr, tooltip="text")
```

```{r}
slr <- ggplot(dat, aes(GDP.per.Capita, Total.Ecological.Footprint)) +
  geom_point(aes(text = Country)) +
  geom_smooth(method= "lm", color = "red", linetype = 1, se=F) +
  geom_smooth(method= "lm", formula = (y ~ x + I(x**2)), color = "blue", linetype = 2, se=F) +
  ggtitle("Simple Linear Regression",
          subtitle = "Model With Quadratic Term does much better") 

ggplotly(slr, tooltip="text")
```


```{r}
ggplot(dat, aes(x = HDI, y = GDP.per.Capita)) +
  geom_point() +
  theme_fivethirtyeight() +
  ggtitle("HDI vs. GDP per Capita") +
   theme(axis.title = element_text(), axis.title.x = element_text())

```

It should be obvious that HDI and GDP are not linearly correlated.



```{r}
multiple1 <- lm(Total.Ecological.Footprint ~  GDP.per.Capita + I(HDI**2) , data = dat)
summary(multiple1)
```

```{r}
ggplot(dat, aes(x = Population..millions., y = Total.Ecological.Footprint)) +
  geom_point() +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), axis.title.x = element_text()) +
  scale_x_log10()
```


## Geographical Features



```{r}

crops <- ggplot(dat, aes(x = Cropland, y = Total.Ecological.Footprint)) +
  geom_point(aes(size = Total.Ecological.Footprint), alpha = 0.4, color = "#e6a526") +  theme_fivethirtyeight() +
  ggtitle("Crop Land") +
   theme(axis.title = element_text(), axis.title.x = element_text(), legend.position="none")


urban <- ggplot(dat, aes(x = Urban.Land, y = Total.Ecological.Footprint)) +
  geom_point(aes(size = Total.Ecological.Footprint), alpha = 0.4, color = "#FF7F50")+
  theme_fivethirtyeight() +
  ggtitle("Urban Land") +
   theme(axis.title = element_text(), axis.title.x = element_text(), legend.position="none")

forest <- ggplot(dat, aes(x = Forest.Land, y = Total.Ecological.Footprint)) +
  geom_point(aes(size = Total.Ecological.Footprint), alpha = 0.4, color = "#56AB6F")+
  theme_fivethirtyeight() +
  ggtitle("Forest Land") +
   theme(axis.title = element_text(), axis.title.x = element_text(), legend.position="none")

fishing <- ggplot(dat, aes(x = Fishing.Water, y = Total.Ecological.Footprint)) +
  geom_point(aes(size = Total.Ecological.Footprint), alpha = 0.4, color = "#504FB1")+
  theme_fivethirtyeight() +
  ggtitle("Fishing Water") +
   theme(axis.title = element_text(), axis.title.x = element_text(), legend.position="none")


cowplot::plot_grid(urban, forest, fishing, crops, ncol = 2, nrow = 2,
                   label_y = "Total Footprint")

```
<br>
We can eyeball that urban land is more important variable than forests and fishing water.



P-value for Urban.Land is quite high, so we will stick to older model.

## Good ol' decision trees

```{r}


d_tree <- rpart(Total.Ecological.Footprint ~ Population..millions. + HDI + GDP.per.Capita + Cropland + Grazing.Land + Forest.Land + Fishing.Water + Urban.Land, data = dat_train_imputed)

rpart.plot(d_tree, main="Decision Tree", fallen.leaves=T, box.palette="GnBu")

predictions_d_tree <- predict(d_tree, newdata = dat_test)
RMSE_d_tree <- sqrt(mean((predictions_d_tree[complete.cases(dat_test)] - dat_test$Total.Ecological.Footprint[complete.cases(dat_test)])**2))
```


## Random Forest Regression

```{r}

indexes <- createDataPartition(dat$Total.Ecological.Footprint, list = F)
dat_train <- dat[indexes, ]
dat_test <- dat[-indexes, ]

dat_train_imputed <- rfImpute(Total.Ecological.Footprint ~Region + Population..millions. + HDI + GDP.per.Capita + Cropland + Grazing.Land + Forest.Land + Fishing.Water + Urban.Land , data = dat_train, iter = 20)


```

```{r}
cv.10.folds <- createMultiFolds(dat_train_imputed, k = 10, times = 2) # cross-validation 10 folds
ctrl.1 <- trainControl(method = "repeatedcv", number = 10, repeats = 3, index = cv.10.folds)

rf_model <- train(Total.Ecological.Footprint ~ Region + Population..millions. + HDI + GDP.per.Capita + Cropland + Grazing.Land + Forest.Land + Fishing.Water + Urban.Land, data = dat_train_imputed, method = "rf", trControl = ctrl.1, tuneLength = 5)

rf_model

```

```{r}
predictions_rf <- predict(rf_model, newdata = dat_test[complete.cases(dat_test), ])


y_real <- dat_test[complete.cases(dat_test), "Total.Ecological.Footprint"]

RMSE_rf <- sqrt(mean((predictions_rf - y_real)**2))
```






